---
title: "Effects of infrastructures for peace"
author: "Maria Benavides"
output: github_document
---


```{r setup, include = FALSE}
library(tidyverse)
library(readxl)
library(here)
library(jcolors)
theme_set(new = theme_minimal())
```


```{r import data, include = FALSE}
# Import FSI downloaded from the website 
fsi_all <- read_excel(here("data/fsi-2006.xlsx"))

for (i in 2007:2020) {           # Use for loop to import all files at once
  fsi_year <- read_excel(here(
    paste(
      "data/fsi-", i, ".xlsx", sep = "")
    )
  )
  if (i %in% c(2019,2020)) {
    fsi_year <- select(fsi_year, -"Change from Previous Year") # Drop extra column for 2019 and 2020
  }
  fsi_all <- rbind(fsi_all, fsi_year) # Combine all data frames into one
}

fsi_all$Year <- as.numeric(str_extract(fsi_all$Year, "[0-9]{4}")) # Modify year variable since it was on a date format 

# Import GPI also downloaded from the website 
gpi_all <- read_xlsx("data/GPI_all.xlsx", sheet = "Overall Scores", skip = 3)

```

## Introduction 

Achieving sustainable peace is one of the most relevant global issues today, especially, since high levels of conflict and violence appeared to be correlated with stagnant development; however, managing to transform violent conflict into stable, nonviolent social and political relationships is a great challenge. 

Recently, a new tool for peacebuilding has caught the attention of scholars and practitioners: **Infrastructures for Peace** (I4P). These can be understood as a dynamic network of skills, capacities, resources, tools and institutions that help build constructive relationships and enhance sustainable resilience of societies against the risks of relapse into violence (Giessmann, 2016).

Considering the above, on this report, I will explore the variations on some positive and negative peace indicators (Fragile State Index and Global Peace Index), on seven countries that have developed government-lead, national peace infrastructures. 

The countries are: 

* Afghanistan
* Cambodia
* Colombia
* Costa Rica
* Ghana
* Kenya
* Yemen

## Analysis of overall scores

```{r data viz fsi, echo=FALSE}
# Plot a graph of the trends of FSI 
fsi_selection <- fsi_all %>%
  select(1,2,4) %>% # Select variables of interest
  filter(
      Country == "Colombia" |
      Country == "Afghanistan" |
      Country == "Cambodia" |
      Country == "Costa Rica" |
      Country == "Ghana" |
      Country == "Kenya" |
      Country == "Yemen" 
    ) %>%
    pivot_wider(
      id_cols = Country,
      names_from = Year, 
      values_from = Total
    ) %>%
    pivot_longer(
      cols = c(2:16), 
      names_to = "Year",
      values_to = "Total"
    ) %>% 
  mutate(Year = as.numeric(Year)) # Set as numeric 
  
# Create a data frame with the years of implementation of infrastructures to use on the graphs
years_impl <- data.frame(Country =  c("Afghanistan", "Cambodia", "Colombia", "Costa Rica", 
                                    "Ghana", "Kenya", "Yemen"
                                      ), 
                          years = c(2010, 2009, 2016, 2009, 2011, 2008, 2011)
                          )  

p <- ggplot(fsi_selection, aes(x = Year, y = Total, group = 1)) + # Plot the graph 
    geom_line() +
    facet_wrap(~Country, 
               scales = "free_x") +
    labs(title="Fragile States Index Variation 2006 - 2020 (2006 score: 100)", 
         x = "Year", 
         y = "Total score") + 
    scale_x_continuous(breaks = c(2006, 2009,2012, 2015, 2018)) 

p + geom_vline(data = years_impl, mapping = aes(xintercept = years)) # Add line of when the infrastructure took place
  
```

These countries adopted at least one of the following infrastructures: national peace councils, government departments, or peace ministries, policy and legislative measures creating I4P, or transitional justice mechanisms. If they were to be effective, we should see an improvement in the **Fragile State Index** overall score in the subsequent years to the implementation; nonetheless, as we can see on the graph, this is not the case for most countries. 

*(Take into consideration that a higher score, means worse conditions).*

Only Colombia and Costa Rica seemed to have improve in terms of state fragility in the years after the implementation of peace infrastructures. The **Yemen** case is also worth mentioning; it presents a drastic worsening overtime, even after the adoption of peace measures.

```{r data viz gpi, echo=FALSE}
# Plot a graph of the trends of GPI
gpi_selection <- gpi_all %>%
  select(-2) %>%   
  filter( # Filter selected countries 
      Country == "Colombia" |
      Country == "Afghanistan" |
      Country == "Cambodia" |
      Country == "Costa Rica" |
      Country == "Ghana" |
      Country == "Kenya" |
      Country == "Yemen" 
    ) %>%
    pivot_longer(
      cols = 2:13,
      names_to = "Year",
      values_to = "Score"
    ) %>%
  mutate(Year = as.numeric(Year)) # Set as numeric 

p <- ggplot(gpi_selection, aes(x = Year, y = Score, group = 1)) + 
    geom_line() +
    facet_wrap(~Country, 
               scales = "free_x") +
    labs(title="Global Peace Index 2008 - 2019", 
         x = "Year", 
         y = "Total score") + 
    scale_x_continuous(breaks = c(2006, 2009,2012, 2015, 2018)) 

p + geom_vline(data = years_impl, mapping = aes(xintercept = years)) # Add line of when the infrastructure took place
  
```

When considering the Global Peace Index, peace infrastructures do not appear to be effective, yet again; there is little variation on the overall scores after the implementation of I4P. 

*(As in the FSI case, a higher value represents less peacefulness)*

It is interesting how there is a high variance (`r var(gpi_selection$Score)`) among the GPI scores; countries like Costa Rica and Ghana are, according to this index, very peaceful, while Yemen and Afghanistan are incredible violent. 

## Analysis of sub-indicators

```{r Sub indices FSI, echo=FALSE}
# Create a data frame with the years of implementation of infrastructures to use on the graphs 
years_post_impl <- data.frame(Country = c("Afghanistan", "Cambodia", "Colombia", "Costa Rica", 
                                        "Ghana", "Kenya", "Yemen"
                                        ),
                              years = c(2011, 2010, 2017, 2010, 2012, 2009, 2012)
                              )

years_post_impl <-years_post_impl %>%
  mutate(Country = as.character(Country)) # Change type to character, because it was set as factor

# Create dataframe with variables of interest

fsi_plot <- fsi_all[0:0,] # Create an empty dataframe


for (n in 1:nrow(years_post_impl)){
  country <- years_post_impl[n,"Country"] # Use loop to filter 
  year <- years_post_impl[n,"years"]
  impl <- filter(fsi_all, Country==country,Year==year)
  
  fsi_plot <- rbind(fsi_plot, impl) # Combine dataframes
} 
  
fsi_plot_2 <- fsi_plot %>% # Modify dataframe for thr graph
  select(-2, -3, -4) %>%
  pivot_longer(
    cols = 2:13, 
    names_to = "indicator",
    values_to = "score"
  ) %>%
  mutate(
    indicator = str_remove(
    indicator, 
    "^[A-Z][0-9]\\:\\s") #Remove extra characters in indicators names
  )

# Plot the subindices 
ggplot(fsi_plot_2, aes(x = score, y = Country, fill = Country)) + 
    geom_bar(stat = "identity") +
    facet_wrap(~indicator, 
               ) +
  scale_fill_jcolors(palette = "pal7") + 
  theme(axis.text.y = element_blank()) + 
  labs(title="Fragile State Index sub-indicators, post I4P implementation", 
         x = "Score", 
         y = "Countries")

```

On the previous graph we see the FSI sub-indicators scores for each country, in the year after they implemented a given peace infrastructure. We can see that **Afghanistan** and **Yemen** perform poorly in all dimensions of state fragility, while **Costa Rica** appears to be a stable nation. 

It also interesting to see that Public Services is the dimension where countries present a lower score. So we could argue that given the fragility of these states, they are unable to provide public services to their citizens. 

## Overall transformation

Although we don’t see much improvement after the implementation of I4P, considering the time frame of the data, some countries have experienced change over time. Based on the FSI, Afghanistan, Colombia, Cambodia and Costa Rica have strengthened their states capacity, as we can see from the variation on the overall score, shown in the following graph. 

*Since a higher score means more fragility, a decrease on it, would represent improvements.*


```{r overall fsi change, echo=FALSE}
# Create dataframe with only the overall scores from 2008 & 2019, to make them comparable 
fsi_change <- fsi_all %>%  
  select(1,2,4) %>%
  filter(Country == "Colombia" |
         Country == "Afghanistan" |
         Country == "Cambodia" |
         Country == "Costa Rica" |
         Country == "Ghana" |
         Country == "Kenya" |
         Country == "Yemen",  
         Year == 2008 |
         Year == 2019
         ) %>%
  rename(
    fsi = Total
  ) %>%
  pivot_wider(
    names_from = Year, 
    values_from = fsi
  ) %>%
  mutate(
    dif = (`2019`-`2008`)/`2008`
  )
  
# Plot the diff of overall score FSI 
ggplot(fsi_change, aes(x = dif, y = Country)) + 
  geom_bar(stat = "identity") +
  scale_x_continuous(labels = scales::percent) +
    labs(title="Fragile State Index variation (2008-2019)", 
         x = "Variation in percentage", 
         y = "Countries")
```

The GPI variation also shows some improvements in terms of peacefulness. But the case of **Yemen** stands out. An increase of over 40% in the overall GPI score, means that violence and conflict have escalated tremendously during the last 11 years. It is also worth noting that **Costa Rica**, although it has presented an increase in the score, it is considered a peaceful country. 


```{r overall gpi change, echo=FALSE}

# Create dataframe with only countries of interest 
gpi_change <- gpi_all %>%  
  select(1,3,14) %>%
  filter(Country == "Colombia" |
         Country == "Afghanistan" |
         Country == "Cambodia" |
         Country == "Costa Rica" |
         Country == "Ghana" |
         Country == "Kenya" |
         Country == "Yemen",  
         ) %>%
  mutate(
    dif = (`2019`-`2008`)/`2008`
  )
  
# Plot the diff of overall score FSI 
ggplot(gpi_change, aes(x = dif, y = Country)) + 
  geom_bar(stat = "identity") +
  scale_x_continuous(labels = scales::percent) +
  labs(title="Global Peace Index variation (2008-2019)", 
         x = "Variation in percentage", 
         y = "Countries")
```

## Conclusions

This exploratory analysis tried to understand how some positive and negative peace indicators changed in countries that implemented infrastructures for peace to eradicate conflict. We saw that the transformation was not in the expected direction, and that in some cases, countries performed worse after implementing these peace policies. 

Given that this report only considered countries that adopted I4P with national coverage, and that were government-lead, we should not conclude that peace infrastructures in general are inefficient. We advise further analysis that include other types of policies and indicators that are more comprehensive. 

